---
title: "Getting Started with xlr8: Creating Styled Excel Reports"
author: "Your Name"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with xlr8}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction

`xlr8` simplifies creating consistently styled Excel reports directly from R. This vignette guides you through populating an Excel template with your data and automatically applying the original formatting.

# Installation

Install the latest version of `xlr8` from GitHub:

```r
# install.packages("devtools")
devtools::install_github("your_github_username/xlr8")
```

# Workflow Overview

`xlr8` automates Excel reporting with three key steps:

1. Prepare an Excel template with embedded metadata tags.
2. Populate the Excel template with your data.
3. Automatically apply styles from the original Excel template.

# Brief Overview of Excel Metadata Tags

The tagging convention used by `xlr8` embeds metadata directly in Excel cells:

- **Single-cell variables** are tagged as:
  ```
  *((var*((variable_name
  ```

- **Tables** are defined as follows:
  - **Table start with first column definition**:
    ```
    *((tbl*((table_name*((col*((first_column_name
    ```
  - **Additional columns** in the table:
    ```
    *((col*((column_name
    ```
  - **Table end marker** placed directly below the table:
    ```
    *((table_end
    ```

For a more detailed explanation on template creation, see the ["Creating Tagged Excel Templates"](creating-tagged-excel-templates.html) vignette.

# Example Usage

First, define your data in R. For templates containing tables, `xlr8` expects your main data frame to include **list-columns containing tibble(s)**. Each list-column corresponds to a table defined in your Excel template metadata.

Here's a clear example of how to structure this:

```{r}
library(xlr8)
library(tibble)

# Single-cell variables
report_date <- as.Date("2025-01-01")
prepared_by <- "Analyst Name"

# Table data prepared separately as a tibble within a list-column
project_table <- tibble(
  project_name = c("Library Expansion", "Community Park"),
  funding_cp = c(500000, 250000),
  funding_cc = c(200000, 150000),
  funding_total = c(700000, 400000),
  opening_date = as.Date(c("2025-06-01", "2026-05-15")),
  date_const = as.Date(c("2024-07-01", "2025-03-20")),
  date_completion = as.Date(c("2025-05-15", "2026-04-30"))
)

# Final dataframe passed to xlr8, including single-cell variables and table data
df <- tibble(
  report_date = report_date,
  prepared_by = prepared_by,
  projects = list(project_table) # The table name matches the metadata tag "projects"
)
```

Now, generate your Excel report using this structured dataframe:

```{r eval=FALSE}
xlr8(
  df = df,
  metadata_path = "templates/project_template.xlsx",
  output_path = "reports/project_report.xlsx"
)
```

`xlr8` will populate the Excel template, automatically applying the original formatting defined by your tagged cells.



# Reviewing the Result

Open your Excel file (`reports/project_report.xlsx`) to confirm data placement and formatting.

# Next Steps

To create or adjust Excel templates, please refer to the dedicated ["Creating Tagged Excel Templates"](creating-tagged-excel-templates.html) vignette.

# Conclusion

You've successfully generated a styled Excel report using `xlr8`. Explore further documentation:

```r
help(package = "xlr8")
```
